Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
zorak4

\section{Discussion}
\label{sec:discussion}

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the accretion rate as a function of black hole mass for the average star formation heating, calculated from equation (\ref{eq:eddr_analytic}) using our results for the total wind heating (Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim 10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{9}M_{\sun}$.  These fall below the maximum thermally stable accretion rate $\dot{M}_{\rm TI}$ (yellow lines), except at the highest black hole masses $\sim 10^{9}M_{\odot}$, for which the older stellar population supplies insufficient stellar wind heating and SN Ia are too rare near the stagnation radius.  In fact, for these highest mass black holes, the maximum accretion rate set by SN Ia blow-out (green lines; eq.~[\ref{eq:eddr_Ia}]) could cap the accretion rate below the thermally-unstable value, provided that the Ia heating efficiency is sufficiently high that $v_{\rm w}^{\rm Ia} \gg \sqrt{2}\sigma \sim 400$ km s$^{-1}$ (see discussion just preceeding eq.~[\ref{eq:eddr_Ia}]).  For lower mass black holes, the maximum Ia accretion rate is significantly higher and cannot prevent thermally unstable flows, i.e. $\dot{M}_{\rm Ia} > \dot{M}_{\rm TI}$.  Finally, also shown for comparison in Figure \ref{fig:bh_growth} (top panel) is the accretion rate onto SgrA* estimated by \citet{Quataert04} and the upper limit onto NGC3115 set by \citealt{Shcherbakov+14}.

@@

The X-ray luminosities of galactic nuclei provide a powerful diagnostic of the SMBH accretion rates and how they vary across different galaxy properties and SMBH masses (e.g.,  \citealt{Ho08}  and references therein). 
appears to contradict the observed behavior of the nuclear X-ray luminosities of quiescent galaxies, which (though showing several orders of magnitude scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha \in [-0.4,-0.2]$ \citep{Zhang+09, Pellegrini10, Gallo+10, Miller+15}.

Figure~\ref{fig:miller} shows our predictions for $\langle L_{X} \rangle =\epsilon_X \epsilon_{\rm bol} \dot{M} c^2$ in Eddington units as a function of black hole mass for core (blue) and cusp (green) galaxies, where $\epsilon_{\rm bol} = L_{X}/L_{\rm bol}= 0.1$ is the assumed bolometric correction appropriate for low luminosity AGN (\citealt{Ho08}).  The left panel is calculated assuming a constant low value for the radiative efficiency $\epsilon_X = 10^{-4}$, typical of those estimated for low luminosity AGN (e.g., \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}).  The luminosity shown in the right
panel instead assumes an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$ that depends on $\dot{M}$ in the way predicted by models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Plotted next to the theoretical curves are the X-ray measurements (stars) and upper limits
(triangles) from \citet{Miller+15} (cf.~\citealt{Gallo+10}), along with the resulting best power-law
fit (red line) and the X-ray detection threshold (gray line).  Purple, orange, and {\bf brown} lines show the maximum accretion rates for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]), as set by SN Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), and as allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]), respectively.

To first order our predicted values of $\langle L_{X}/L_{\rm edd}\rangle$ for the average galactic star formation histories overlap with the data reasonably well.  Our prediction that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  However, better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude can also be understood within a steady-accretion scenario due to the strong sensitivity of the accretion rate on the efficacy of gas heating, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the equally sensitive dependence of the heating rate on stochastic intermittency in the star formation history (Fig.~NICK).

@@

\subsection{Nuclear Star Clusters}
Do nuclear star clusters in other galaxies look like scaled up/down
versions of the one in our own galaxy? If so this would imply a radial
gradient in the average of the stellar population which would imply a
heating which would be dependent on radius.

\subsection{SFR-AGN connection}
In fact it seems like the star formation rate may trace AGN
episodes. See e.g. http://arxiv.org/pdf/1501.07602.pdf.

\subsection{Type II Supernovae}
Type II supernovae should provide some heating as well for younger
stellar populations...In fact at least on large scales they may be the
dominant source of heating on timescales of 10-40 Myr.

@@

Apologies: we assume that average star formation histories correspond to those experienced in the nuclear region.

>>>>

\section{Implications and Discussion}
\label{sec:discussion}

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the accretion rate as a function of black hole mass for the average star formation heating, calculated from equation (\ref{eq:eddr_analytic}) using our results for the total wind heating (Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim 10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{9}M_{\sun}$.  These fall below the maximum thermally stable accretion rate $\dot{M}_{\rm TI}$ (yellow lines), except at the highest black hole masses $\sim 10^{9}M_{\odot}$, for which the older stellar population supplies insufficient stellar wind heating and SN Ia are too rare near the stagnation radius.  In fact, for these highest mass black holes, the maximum accretion rate set by SN Ia blow-out (green lines; eq.~[\ref{eq:eddr_Ia}]) could cap the accretion rate below the thermally-unstable value, provided that the Ia heating efficiency is sufficiently high that $v_{\rm w}^{\rm Ia} \gg \sqrt{2}\sigma \sim 400$ km s$^{-1}$ (see discussion just preceeding eq.~[\ref{eq:eddr_Ia}]).  This is only possible if the Ia heating efficiency is reasonably high $\epsilon_{\rm Ia} > 0.1$ than normally assumed (\citealt{Sharma+14}).  For lower mass black holes, the maximum Ia accretion rate is significantly higher and cannot prevent thermally unstable flows, i.e. $\dot{M}_{\rm Ia} > \dot{M}_{\rm TI}$.

Finally, also shown for comparison in Figure \ref{fig:bh_growth} (top panel) is the accretion rate onto SgrA* estimated by \citet{Quataert:2004a} and the upper limit onto NGC3115 set by \citealt{ShcherbakovWong+:2014a}.

@@

The X-ray luminosities of galactic nuclei provide a powerful diagnostic of the SMBH accretion rates and how they vary across different galaxy properties and SMBH masses (e.g.,  \citealt{Ho08}  and references therein).  Figure~\ref{fig:miller} shows our predictions for $\langle L_{X} \rangle =\epsilon_X \epsilon_{\rm bol} \dot{M} c^2$ in Eddington units as a function of black hole mass, where $\epsilon_{\rm bol} = L_{X}/L_{\rm bol}= 0.1$ is the assumed bolometric correction appropriate for low luminosity AGN (\citealt{Ho08}).  We have averaged the accretion rate used in calculating $\langle L_X \rangle$ between the values for galaxies with core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) stellar density profiles, by fiting the average value of the inner density slope $\Gamma$ as a function of black hole mass from the \citet{LauerFaber+:2007a} galaxy sample.

The left panel of Figure \ref{fig:miller} is calculated assuming a constant low value for the radiative efficiency $\epsilon_X = 10^{-4}$, typical of those estimated for low luminosity AGN (e.g., \citealt{Ho:2009a}) and SgrA* (e.g., \citealt{Quataert+99}).  The luminosity shown in the right panel instead assumes an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$ that depends on $\dot{M}$ in the way predicted by models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Plotted next to the theoretical curves are the X-ray measurements (green stars) and upper limits (gray triangles) compiled by \citet{Miller+15} (cf.~\citealt{Gallo+10}). A green line shows the resulting best power-law fit to the average X-ray luminosity, which is given by $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ from \citet{Miller+15}, with $\alpha = XXX$  (see also \citep{Zhang+09, Pellegrini10, Gallo+10} for similar conclusions).  Yellow, teal, and red lines show the maximum accretion rates, respectively, for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]), as set by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]), and as allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]).

To first order our predicted average values of $\langle L_{X}/L_{\rm edd}\rangle$ increases only weakly with black hole mass.

for the average galactic star formation histories overlap with the data reasonably well.  Our prediction that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  However, better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude can also be understood within a steady-accretion scenario due to the strong sensitivity of the accretion rate on the efficacy of gas heating, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the equally sensitive dependence of the heating rate on stochastic intermittency in the star formation history (Fig.~NICK).

@@

\subsection{Nuclear Star Clusters}

@@

Apologies: 
\begin{itemize}
\item{we assume that average star formation histories correspond to those experienced in the nuclear region.  }

\item{
Do nuclear star clusters in other galaxies look like scaled up/down versions of the one in our own galaxy? If so this would imply a radial gradient in the average of the stellar population which would imply a heating which would be dependent on radius.
}

\end{itemize}
